---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

In this document I am going to make the Peer Assessment of the Course of Coursera Reproducible Research.


## Loading and preprocessing the data
```{r loading, echo=TRUE}
knitr::opts_chunk$set(tidy=FALSE, fig.path='figures/')

file <- unzip("activity.zip")
data <- read.csv(file)

head(data)
```


## What is mean total number of steps taken per day?
```{r steps, echo=TRUE, fig.height=6}
stepsbydate <- aggregate(steps ~ date, data = data, FUN=sum)
barplot(stepsbydate$steps, main ="Number of steps taken per day", names.arg=stepsbydate$date)
```

Now we are going to calculate the mean and the median of steps per day
```{r meanandmedian}
meanSteps <- round(mean(stepsbydate$steps),2)
medianSteps <- round(quantile(stepsbydate$steps, probs = 0.5),2)
```
The mean steps per day is `r sprintf("%.0f",meanSteps)` and the median steps is `r sprintf("%.0f",medianSteps)`

## What is the average daily activity pattern?
```{r interval, echo=TRUE, fig.height=6}
stepsbyinterval <- aggregate(steps ~ interval, data = data, FUN=mean)
plot(x=stepsbyinterval$interval, stepsbyinterval$steps, main ="Daily activity pattern", type="l")
```

```{r}
si <- stepsbyinterval[order(stepsbyinterval$steps, decreasing = TRUE),]
maximum <- si[1,1]
maxvalue <- si[1,2]
```
The interval labeled as `r maximum` is where there are the maximum number of steps average a day(`r maxvalue`).


## Imputing missing values
```{r imputingincompletes}
#Number of cases with NA.
incompletes <- sum(!complete.cases(data))
``` 

The number of missing values in the dataset is `r incompletes` so we are going to replace de NA values for the mean of that interval.

```{r stepscompletes}
#Round the mean of the steps
stepsbyinterval$steps <- floor(stepsbyinterval$steps)
datacomplete <- merge(data, stepsbyinterval, by="interval", suffixes=c("",".y"))
datacomplete$steps[is.na(datacomplete$steps)] <- datacomplete$steps.y[is.na(datacomplete$steps)]
datacomplete <- datacomplete[,c(1:3)]
stepscompletebydate <- aggregate(steps ~ date, data = datacomplete, FUN=sum)
barplot(stepscompletebydate$steps, main ="Number of steps taken per day with NA filled", names.arg=stepscompletebydate$date)
```

```{r}
meanStepsCom <- round(mean(stepscompletebydate$steps),2)
medianStepsCom <- round(quantile(stepscompletebydate$steps, probs = 0.5),2)
```
The mean steps per day is `r sprintf("%.0f",meanStepsCom)` and the median steps is `r sprintf("%.0f",medianStepsCom)`. Which is not more differente between mean (`r sprintf("%.0f", meanSteps)`) and median(`r sprintf("%.0f",medianSteps)`) without replacing NA.



## Are there differences in activity patterns between weekdays and weekends?

```{r}
day <- function(date){
  if(weekdays(as.Date(date)) %in% c("sábado","domingo")){
    "weekend"
  }else{
    "weekday"
  }
}
datacomplete$day <- as.factor(sapply(datacomplete$date, day))
```

Watching the difference between both plots, we see that there are diference between weekdays and weekends. In weekends there are more steps at every interval in the day. But in weekdays there are a big ammount of steps at 830 interval (before starting jobs).


```{r graphicsday,fig.height=9}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    steps.type <- aggregate(steps ~ interval, data = datacomplete, subset = datacomplete$day == type, FUN = mean)
    plot(steps.type, type = "l", main = type)
}

```








